Supporting Information for\\ ``A weakly structured stem for human origins in Africa''

### 2022-12-19 paper.tex
add Godambe table, unfinished

\usepackage{authblk}
\renewcommand{\Affilfont}{\normalsize}
\renewcommand{\Authfont}{\normalsize}

@@

We quantify migration rates among populations after their divergence $\approx120ka$.
Migration, prior to the Bantu expansion, between the ancestors of the Nama and other groups
is an order of magnitude weaker than that observed between between western and eastern Africans (Table \ref{tab:migration-rates}).
All models infer relatively high gene flow between Ethiopia and western Africa
($m\approx2\times10^{-4},$ the constant proportion of immigrant lineages per generation over 60ka).
We further find that Back-to-Africa gene flow at the beginning of the Holocene primarily
affected the ancestors of the Ethiopian agricultural populations \citep{Molinaro2019-rf},
comprising over half of their genetic ancestry, estimated to be 64--65\%.
We observe significant gene flow from the Amhara and Oromo into the Nama, a signal which is likely a proxy for the movement of
eastern African caprid and cattle pastoralists \citep{Henn2008-xo,Breton2014-xb}, here estimated to constitute a 25\% ancestry
contribution 2,000 ya. While this gene flow is not apparent from the ADMIXTURE plot (Figure~\ref{fig:diversity}),
the ancestry is likely grouped into the Khoe-San component which has drifted appreciably from its ancestral eastern African source.
Colonial period admixture from Europeans into the Nama
was estimated at 15\%, similar to proportions suggested by \texttt{ADMIXTURE}
(Figure~\ref{fig:diversity}).

>>>>

\usepackage{authblk}

\usepackage{lineno}
\linenumbers
\renewcommand{\Affilfont}{\normalsize}
\renewcommand{\Authfont}{\normalsize}

@@

We quantify migration rates among populations after their divergence
$\approx120ka$. Prior to the Bantu expansion, migration between the ancestors
of the Nama and other groups is an order of magnitude weaker than that observed
between between western and eastern Africans (Table \ref{tab:migration-rates}).
All models infer relatively high gene flow between Ethiopia and western Africa
($m\approx2\times10^{-4}$, the constant proportion of immigrant lineages per
generation over 60ka). We further find that Back-to-Africa gene flow at the
beginning of the Holocene primarily affected the ancestors of the Ethiopian
agricultural populations \citep{Molinaro2019-rf}, comprising over half of their
genetic ancestry, estimated to be 64--65\%.  We observe significant gene flow
from the Amhara and Oromo into the Nama, a signal which is likely a proxy for
the movement of eastern African caprid and cattle pastoralists
\citep{Henn2008-xo,Breton2014-xb}, here estimated to constitute a 25\% ancestry
contribution 2,000 ya. While this gene flow is not apparent from the ADMIXTURE
plot (Figure~\ref{fig:diversity}), the ancestry is likely grouped into the
Khoe-San component which has drifted appreciably from its ancestral eastern
African source.  Colonial period admixture from Europeans into the Nama was
estimated at 15\%, similar to proportions suggested by \texttt{ADMIXTURE}
(Figure~\ref{fig:diversity}).

### 2022-12-19 supplement.tex
add Godambe table, unfinished

\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}

@@

@@

@@

@@

>>>>

\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{lineno}
\linenumbers

@@

\aprcomment{TODO: add (here?) our reasoning and validation of the choice of
60/50ka for fixed divergence times.}

@@

\aprcomment{To ensure that our confidence intervals computed using the approach
    from \citet{Coffman2016-yq} do not underestimate true CIs, we also
    performed block bootstrap resampling of our data, 200 times, and refit each
    model to the bootstrapped datasets. From these, we evaluated the
distribution of inferred parameters, which should be similar to or narrower
than the CIs estimated from the GIM approach. See
Section~\ref{sec:bootstrap-CIs}.}

@@

\subsection{Evaluation of confidence intervals}
\label{sec:bootstrap-CIs}

\aprcomment{Here, point to tables for CIs, or SEs from Godambe vs reinferred
models from bootstrapped data. See Table~\ref{tab:supp-confidence-intervals}}

\subsection{Allowing for additional migrations}

\aprcomment{Here, addressing major comment 3 from reviewer 4, migration between
Nama and Mene and Nama and GBR.}

\subsection{Replacing the reference population for statistics normalization}

\aprcomment{Minor comment 3 from reviewer 4.}

@@

\begin{table}[ht]
    \caption{
        \label{tab:supp-confidence-intervals}
        \textbf{GIM-computed confidence intervals are not underestimated.}
        CIs computed with the Godambe information matrix (GIM) are larger than
        those estimated by creating block bootstrap-replicate datasets and
        refitting each model (200 times) to these datasets. Standard deviations
        for each inferred parameter were then computed from these 200 model fits.
        Parameters are as labeled and described in the above tables.
        Blank standard errors indicate that parameter was not relevant
        to the given model.
        \aprcomment{Fill in after perturbing starting conditions}
    }
    \centering
    \begin{tabular}[t]{l|cc|cc|cc|cc}
        \toprule
        & \multicolumn{2}{c|}{Single origin} & \multicolumn{2}{c|}{Cont. mig.}
        & \multicolumn{2}{c|}{Merger w/o mig.} & \multicolumn{2}{c}{Merger w/mig.} \\
        Parameter & GIM & BS & GIM & BS & GIM & BS & GIM & BS \\
        \midrule
        $N_e$ & 403 & 127 & 1777 & 951 & 326 &  & 1369 &  \\
        $N_{MH}$ & 529 & 158 & & & & & & \\
        $N_{stem1}$ & & & 1612 & 406 & 76 &  & 838 &  \\
        $N_{stem2}$ & & & 2488 & 1001 & 1149 &  & 6668 &  \\
        $N_{Nama_0}$ & 370 & 79 & 2989 & 512 & 384 &  & 1514 &  \\
        $N_{Nama_F}$ & 9 & 2 & 54 & 4 & 7 &  & 31 &  \\
        $N_{MSL_0}$ & 769 & 190 & 2479 & 285 & 322 &  & 1165 &  \\
        $N_{MSL_F}$ & 606 & 185 & 6628 & 1250 & 987 &  & 4332 &  \\
        $N_{EA}$ & 273 & 49 & 1841 & 138 & 246 &  & 1628 &  \\
        $N_{Gumuz_F}$ & 131 & 56 & 915 & 80 & 102 &  & 337 &  \\
        $N_{EP}$ & 491 & 106 & 3246 & 168 & 408 &  & 1684 &  \\
        $N_{GBR_0}$ & 33 & 7 & 231 & 17 & 29 &  & 122 &  \\
        $N_{GBR_F}$ & 507 & 117 & 2889 & 295 & 334 &  & 1678 &  \\
        $N_{Neand}$ & 105 & 88 & 591 & 97 & 91 &  & 235 &  \\
        $T_{stems}$ & & & 391E3 & 67E3 & 27E3 &  & 426E3 &  \\
        $T_{stem1}$ & & & & & 20E3 &  & 166E3 &  \\
        $m_{Nama-MSL}$ & 1.6E-6 & 1.2E-6 & 3.7E-6 & 2.7E-6 & 1.1E-6 &  & 4.0E-6 &  \\
        $m_{Nama-EA}$ & 2.0E-6 & 1.3E-6 & 1.0E-5 & 1.8E-6 & 1.3E-6 &  & 9.1E-6 &  \\
        $m_{MSL-EA}$ & 7.6E-6 & 1.1E-6 & 5.3E-5 & 3.0E-6 & 5.5E-6 &  & 2.6E-5 &  \\
        $m_{EA-GBR}$ & 2.4E-6 & 1.1E-6 & 1.0E-5 & 1.9E-6 & 1.5E-6 &  & 5.5E-6 &  \\
        $m_{EA-EA}$ & 1.3E-5 & 7.5E-6 & 8.4E-5 & 1.4E-5 & 1.3E-5 &  & 5.4E-5 &  \\
        $f_{GBR \rightarrow EP}$ & 0.0039 & 0.0036 & 0.0037 & 0.0071 & 0.0037 &  & 0.0075 &  \\
        $f_{EP \rightarrow Nama}$ & 0.0039 & 0.0040 & 0.0043 & 0.0072 & 0.0042 &  & 0.0049 &  \\
        $f_{GBR \rightarrow Nama}$ & 0.0019 & 0.0020 & 0.0021 & 0.0030 & 0.0021 & & 0.0031 &  \\
        $m_{stems}$ & & & 1.1E-5 & 7.7E-6 & & & 8.7E-5 &  \\
        $m_{stem2-Nama}$ & & & 1.6E-5 & 1.1E-5 & & & & \\
        $m_{stem2-MSL}$ & & & 4.2E-5 & 1.7E-5 & & & & \\
        $m_{stem2-EA}$ & & & 9.0E-6 & 5.4E-6 & & & & \\
        $T_{Nama}$ & 2525 & 1188 & 18E3 & 6400 & 8253 &  & 28E3 &  \\
        $f_{stem 2 \rightarrow Nama}$ & & & & & 0.0086 &  & 0.067 &  \\
        $T_{EA}$ & & & & & 3648 &  & 8865 &  \\
        $f_{stem 2 \rightarrow EA}$ & & & & & 0.0074 &  & 0.059 &  \\
        $T_{MSL}$ & & & & & 570 &  & 641 &  \\
        $f_{stem 2 \rightarrow MSL}$ & & & & & 0.0036 &  & 0.0085 &  \\
        \bottomrule
    \end{tabular}
\end{table}
